---
title: "RWorksheet_#4"
author: "Nicole Duero"
date: "2023-10-25"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. The table below shows the data about shoe size and height. Create a data frame. a. Describe the data.
```{r}
Shoe_size<-c(6.5,9.0,8.5,8.5,10.5,7.0,9.5,9.0,13.0,7.5,10.5,8.5,12.0,10.5,13.0,11.5,8.5,5.0,10.0,6.5,7.5,8.5,10.5,8.5,10.5,11.0,9.0,13.0)
Height<-c(66.0,68.0,64.5,65.0,70.0,64.0,70.0,71.0,72.0,64.0,74.5,67.0,71.0,71.0,77.0,72.0,59.0,62.0,72.0,66.0,64.0,67.0,73.0,69.0,72.0,70.0,69.0,70.0)
HouseholdData<-data.frame(Shoe_size,Height)
HouseholdData

Gender<-c("F","F","F","F","M","F","F","F","M","F","M","F","M","M","M","M","F","F","M","F","F","M","M","F","M","M","M","M")
HouseholdDataNew<-cbind(HouseholdData, Gender)
HouseholdDataNew
```
ANSWER: The data provided contains information about individuals' shoe sizes, heights, and genders. Additionally, there are 28 observations in the dataset.

b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.
```{r}
males<-subset(HouseholdDataNew, Gender=="M")
males
females<-subset(HouseholdDataNew, Gender=="F")
females
```
c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
mShoe<-mean(HouseholdDataNew$Shoe_size)
mShoe
mHeight<-mean(HouseholdDataNew$Height)
mHeight
```
d. Is there a relationship between shoe size and height? Why?
```{r}
RelShoeHeight<- cor(HouseholdDataNew$Shoe_size, HouseholdDataNew$Height)
RelShoeHeight
```
ANSWER: Yes, as it indicates a strong positive linear relationship between shoe size and height. This means that as shoe size increases, there is a tendency for height to also increase.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
```{r}
months_vector<-c("March","April","January","November","January","September","October","September","November","August","January","November","November","February","May","August","July","December","August","August","September","November","February","April")
factor_months_vector<-factor(months_vector)
factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
summary(factor_months_vector)
```
ANSWER: Yes, they are both equally useful. 
If you view months as text labels, use `months_vector` as it's a character vector.
For statistical operations or categorizing months, `factor_months_vector` is more suitable.

#4. Create a vector and factor for the table below.
```{r}
Direction<-c("East", "West","North")
Frequency<-c(1,4,3)
new_order_data<-factor(Direction, levels=c("East", "West","North"))
print(new_order_data)
new_order_data1<-factor(Frequency, levels=c(1,4,3))
print(new_order_data1)
```
5. Enter the data below in Excel with file name = import_march.csv
a. Import the excel file into the Environment Pane using read.table() function. Write the code.b. View the dataset. Write the R scripts and its result.
```{r}
read_importMarch<-read.table(file="C://Users//User//Documents//Rstudio Files//Worksheet#4//import_march.csv", header=TRUE, sep=",")
read_importMarch
```
6. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string “The number selected is beyond the range of 1 to 50”. If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.
```{r}
UserNum<-readline(prompt="Enter a number from 1 to 50: ")
if(UserNum == 20){
  print("TRUE")
}else if(UserNum<=50 && UserNum>=1){
   cat("The input number is:",UserNum) 
}else{
    print("The number selected is beyond range of 1 to 50.")
}
```
7.a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack. Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.
```{r}
min_bills<-function(){
 price<-as.integer(readline(prompt="Price of snack(a random number divisible by 50):"))

 if (is.na(price)|| price %% 50 !=0){
  cat("Invalid. Should be divisible by 50.")
return()
}
  num_bills<-0
  bill_type<-c(1000,500,200,100,50)
  
  for(bill in bill_type){
    num_bills<-num_bills + (price %/% bill)
    price<-price %%bill
  }
  
cat("Minimum number of bills needed:", num_bills,"\n")
}
min_bills()
```
```{r}
#8. a. Create a dataframe from the above table. Write the R codes and its output.
Name<-c("Annie", "Thea", "Steve", "Hanna")
Grade1<-c(85, 65, 75, 95)
Grade2<-c(65, 75, 55, 75)
Grade3<-c(85, 90, 80, 100)
Grade4<-c(100, 90, 85, 90)

StudentsScore<-data.frame(Name, Grade1, Grade2, Grade3, Grade4)
StudentsScore

#b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output. Example Output: Annie’s average grade this semester is 88.75.
students_above_90<-FALSE
for(g in 1:length(Name)){
  AveScore<-c((Grade1)[g]+(Grade2)[g]+(Grade3)[g]+(Grade4)[g])/4
  if(AveScore>90){
  cat(paste(Name[g], "'s average grade this semester is", round(AveScore,2),".\n"))
    students_above_90<-TRUE
  }
}
if(!students_above_90){
  print("No students has reached above 90 average math score.")
}

#c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests. Example output: The nth test was difficult.
for (test_num in 1:4){
  AveScore2<-Grade1 + Grade2 + Grade3 + Grade4
  AveScore2<-AveScore2/4
  if (AveScore2[test_num]<80){
    cat("The", test_num, "test was difficult.\n")
  }
}

#d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points. Example Output: Annie’s highest grade this semester is 95.
for (h in 1:length(Name)){
  highGrade<-Grade1[h]
  
  if (Grade2[h]>highGrade){
    highGrade<-Grade2[h]
  }
  if (Grade3[h]>highGrade){
    highGrade<-Grade3[h]
  }
  if (Grade4[h]>highGrade){
    highGrade<-Grade4[h]
  }
  
  if (highGrade>90){
    cat(paste(Name[h], "'s highest grade this semester is", highGrade, ".\n"))
  }
}
```


