---
title: "RWorksheet_DUERO#4b"
author: "Nicole Duero"
date: "2023-11-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
```
Using Loop Function
for() loop

NUMBER 1
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint: Use abs() function to get the absolute valu
```{r}
vectorA<-c(1,2,3,4,5)
matrixA<-matrix(0, nrow=5, ncol=5)
matrixA

for(i in 1:5){
  for(j in 1:5){
   matrixA[i,j]<- abs(vectorA[i] - vectorA[j])
  }
}

matrixA

```
NUMBER 2
2. Print the string "*" using for() function. The output should be the same as shown in Figure 2. 
```{r}
library(RColorBrewer)
asterisk<-5

for(i in 1: asterisk){
  for(j in 1:i){
    cat("* ")
  }
  cat("\n")
}

```
NUMBER 3
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
userInput<-as.integer(readline(prompt="Input a number of terms: "))
a<-0
b<-1
cat("Fibonacci Sequence: ", a, b)

repeat{
  c<-a+b
  if(c>500){
    break
  }
  cat(", ", c)
  a<-b
  b<-c
}
```
NUMBER 4
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result. 

```{r}
library(readr)
csv.file<-"HouseholdData.csv"
HouseholdData<-read.csv("HouseholdData.csv")
HouseholdData
head(HouseholdData,6)
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
HD1<-subset(HouseholdData, Gender=="M")
HD1
male_gender<-nrow(HD1)
male_gender
#14 observations for Male.

HD2<-subset(HouseholdData, Gender=="F")
HD2
female_gender<-nrow(HD2)
female_gender
#14 observations for Female.
```
c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r}
count<-c(male_gender, female_gender)
Gender<-c("Male","Female")
barplot(count,
        names.arg=Gender, 
        main="Number of Males of Females in Household Data",
        xlab="Gender",
        ylab="Count",
        col=c("blue", "pink")
        )
legend("topright",
       legend=Gender,
       fill=c("blue", "pink")
       )
```
NUMBER 5
a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
```{r}
monthlyIncome<-c(60, 10, 5, 25)

monthLabels<-round(monthlyIncome/sum(monthlyIncome)*100,1)
monthLabels<-paste(monthLabels,"%", sep="")

pie(monthlyIncome, main="Monthly Income of Dela Cruz Family", col=rainbow(length(monthlyIncome)), labels=monthLabels, cex=0.8)
legend(1.5, 0.5, c("Food", "Electricity", "Savings", "Misscelanous"), cex=0.8, fill=rainbow(length(monthlyIncome)))
```
NUMBER 6
6. Use the iris dataset.
data(iris)

a. Check for the structure of the dataset using the str() function. Describe what you
have seen in the output.

b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?

c.Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.
```{r}
data(iris)
str(iris)

mean_iris<-colMeans(iris[,c("Sepal.Length", "Sepal.Width","Petal.Length", "Petal.Width")])
mean_iris

pie(table(iris$Species),
    main="Species Distribution",
    labels= levels(iris$Species),
    col=c("red", "orange", "blue"))
legend("topleft",
       legend=levels(iris$Species),
       fill=c("red", "orange", "blue"),
       title="Species Distribution")
```
d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r}
setosaSix<- tail(subset(iris, Species == "setosa"), n = 6)
versicolorSix <- tail(subset(iris, Species == "versicolor"), n = 6)
virginicaSix<- tail(subset(iris, Species == "virginica"), n = 6)
setosaSix

versicolorSix

virginicaSix

```
e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
     pch=as.integer(iris$Species),
     col=as.integer(iris$Species),
     main="Iris Data Set",
     sub="Sepal Width and Sepal Length",
     xlab="Sepal Length",
     ylab="Sepal Width",
     cex=1.5, 
     lwd=1.5)
```
Hint: Need to convert to factors the species to store categorical variables.
f. Interpret the result
```{r}
as.factor(iris$Species)
```
NUMBER 7
7. Import the alexa-file.xlsx. Check on the variations. Notice that there are extra whitespaces among black variants (Black Dot, Black Plus, Black Show, Black
Spot). Also on the white variants (White Dot, White Plus, White Show, White
Spot)
a. Rename the white and black variants by using gsub() function. 
```{r}
library(readxl)
alexa_file<-read_excel("alexa_file.xlsx")

alexaFileVariation <- gsub("Black Plus", "Black Plus", alexa_file$variation)
alexa_file$variation <- gsub("Black Show", "Black Show", alexa_file$variation)
alexa_file$variation <- gsub("Black Spot", "Black Spot", alexa_file$variation)
alexa_file$variation <- gsub("Black Dot", "Black Dot", alexa_file$variation)
alexa_file$variation <- gsub("White Dot", "White Dot", alexa_file$variation)
alexa_file$variation <- gsub("White Plus", "White Plus", alexa_file$variation)
alexa_file$variation <- gsub("White Show", "White Show", alexa_file$variation)
alexa_file$variation <- gsub("White Spot", "White Spot", alexa_file$variation)

knitr::include_graphics("/cloud/project/Worksheet_DUERO#4/Worksheet_#4b")
```
b. Get the total number of each variations and save it into another object. Save the
object as variations.RData. Write the R scripts. What is its result?
Hint: Use the dplyr package. Make sure to install it before loading the package.
```{r}
library(dplyr)

save(alexa_file, file = "variations.RData")
load("variations.RData")
alexaVar<- alexa_file%>%count(alexa_file$variation)
alexaVar
```
c. From the variations.RData, create a barplot(). Complete the details of the chart
which include the title, color, labels of each bar.
```{r}
barplot(
  height=alexaVar$n,
  names.arg=alexaVar$`alexa_file$variation`,
  col="lightgreen",
  main="Variation of Alexa",
  las=2,
  cex.names=0.58
)
```
d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by
side. Complete the details of the chart.
```{r}
par(mfrow = c(1, 2))
black_variants <- alexaVar[1:5,]
white_variants <- alexaVar[12:16,]

barplot(
height = black_variants$n,
names.arg = black_variants$`alexa_file$variation`,
main = "Black Variants",
col = c("red", "orange", "yellow", "green", "blue"),
xlab = 'Total Numbers',
ylab = 'Frequency',
cex.names = 0.35,
)

barplot(
height = white_variants$n,
names.arg = white_variants$`alexa_file$variation`,
main = "White Variants",
col = c("violet", "pink", "lightblue", "gray", "black"),
xlab = 'Total Numbers',
ylab = 'Frequency',
cex.names = 0.35,
)
```